Waveform and Passive Beamforming Design for Intelligent Reflecting Surface-Aided Wireless Information and Power Transfer

### 2020-07-03 slides.tex
:pencil: System model and problem formulation

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{cuted}
\usepackage{hyperref}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{subcaption}

@@

		\begin{frame}{WPT by RF waves}
			\textbf{Energy flow}: DC $\to$ RF $\to$ RF $\to$ DC
			\begin{figure}
				\centering
				\includegraphics[width=\textwidth]{assets/wpt.eps}
			\end{figure}
			\textbf{Pros}:
			\begin{itemize}
				\item long range (up to hundreds of \si{\m}) with NLoS support
				\item compact receiver (few \si{\cm}), easy integration
				\item suitable for mobile devices
			\end{itemize}
			\textbf{Cons}:
			\begin{itemize}
				\item low power level (\si{\uW} -- \si{\mW})
				\item low energy harvesting efficiency (40\% at 100 \si{\uW}, 20\% at 10 \si{\uW})
			\end{itemize}
			\blfootnote{Figure from \cite{Clerckx2018a}}
		\end{frame}
	\end{subsection}

@@

		\begin{frame}{Harvester model}
			RF-to-DC conversion requires \textbf{rectenna} (receive antenna + rectifier), whose behavior is dominated by diode I-V characteristics.
			\begin{figure}
				\centering
				\includegraphics[width=0.8\textwidth]{assets/rectenna_circuit.eps}
				\caption{Rectenna equivalent circuit and a single diode rectifier \cite{Clerckx2018a}}
			\end{figure}
			Consider small-signal model and truncate its Taylor expansion to the $n_0$-th order:
			\begin{itemize}
				\item diode linear model (${n_o} = 2$): output power is proportional to input power
				\item \alert{diode nonlinear model} (${n_o} > 2$): significant contribution from high-order terms
			\end{itemize}
		\end{frame}

@@

		\begin{frame}{Why IRS-aided SWIPT?}
			\begin{itemize}
				\item both aim at improving spectral/energy efficiency
				\item enhanced channel boosts received power to benefit from harvester nonlinearity
				\item extra links increase system diversity and stability, which is essential for SWIPT
				\item SWIPT can potentially support low-power IRS
			\end{itemize}
		\end{frame}
	\end{subsection}
\end{section}

>>>>

\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{cuted}
\usepackage{hyperref}
\usepackage{lmodern}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{subcaption}
\usepackage[short]{optidef}

@@

		\begin{frame}{WPT by RF waves}
			\textbf{Energy flow}: DC $\to$ \alert{RF $\to$ RF $\to$ DC}
			\begin{figure}
				\centering
				\includegraphics[width=\textwidth]{assets/wpt.eps}
			\end{figure}
			\textbf{Pros}:
			\begin{itemize}
				\item long range (up to hundreds of \si{\m}) with NLoS support
				\item compact receiver (few \si{\cm}), easy integration
				\item suitable for mobile devices
			\end{itemize}
			\textbf{Cons}:
			\begin{itemize}
				\item low power level (\si{\uW} -- \si{\mW})
				\item low energy harvesting efficiency (40\% at 100 \si{\uW}, 20\% at 10 \si{\uW})
			\end{itemize}
			\blfootnote{Figure from \cite{Clerckx2018a}}
		\end{frame}
	\end{subsection}

@@

		\begin{frame}{Harvester model}
			RF-to-DC conversion requires \textbf{rectenna} (receive antenna + rectifier), whose behavior is dominated by diode I-V characteristics.
			\begin{figure}
				\centering
				\includegraphics[width=0.8\textwidth]{assets/rectenna_circuit.eps}
				\caption{Rectenna equivalent circuit and a single diode rectifier \cite{Clerckx2018a}}
			\end{figure}
			Consider small-signal model and truncate its Taylor expansion to the $n_0$-th order:
			\begin{itemize}
				\item diode linear model (${n_o} = 2$): output power is proportional to input power
				\item \alert{diode nonlinear model} (${n_o} > 2$): contribution from high-order terms
			\end{itemize}
			\blfootnote{Figure from \cite{Clerckx2019}}
		\end{frame}

@@

		\begin{frame}{Why IRS-aided SWIPT?}
			\begin{itemize}
				\item both aim at improving spectral/energy efficiency
				\item enhanced channel boosts received power to benefit from harvester nonlinearity
				\item extra links increase system diversity and stability, which is essential for SWIPT
				\item SWIPT can potentially support low-power IRS
			\end{itemize}
		\end{frame}
	\end{subsection}

	\begin{subsection}{Previous research}
		\begin{frame}{Existing works}
			Separated information decoder and energy harvester:
			\begin{itemize}
				\item \cite{Wu2019b} investigated a MISO system with energy interference and proved at most one energy beam is required to maximize the WSP subject to SINR constraints.
				\item \cite{Tang2019} considered fairness issue for a MISO system and maximized the minimum output power based on perfect energy interference cancellation.
				\item \cite{Pan2019a} transformed WSR maximization of MIMO SWIPT to WMMSE problem then solved by BCD with low-complexity iterative algorithms.
				\item \cite{Wu2019c} proposed a penalty-based algorithm to minimize transmit power of MISO system subject to SINR constraints, whose inner layer updates precoders, phase shifts and auxiliary variables by BCD while the outer layer updates the penalty coefficients.
			\end{itemize}
		\end{frame}

		\begin{frame}{Outcomes and opportunities}
			\textbf{Outcomes}:
			\begin{itemize}
				\item IRS brings a significant gain to SWIPT
				\item performance vary significantly under different configuration (distances, number of reflectors, transmit power, SNR, etc.)
				\item dedicated energy beam is necessary to boost the harvested power
				\item LoS links can further increase the harvested power, since rank-deficient channels are highly correlated and a single energy stream can provide large output power for multiple harvesters
			\end{itemize}
			\vspace{1em}
			\textbf{Opportunities}:
			\begin{itemize}
				\item no works on co-located information decoder and energy harvester
				\item no works on IRS-aided OFDM SWIPT
				\item no works consider harvester nonlinearity
				\item no works investigate Rate-Energy (R-E) tradeoff
			\end{itemize}
		\end{frame}
	\end{subsection}
\end{section}

\begin{section}{System model}
	\begin{subsection}{Signal and channel}
		\begin{frame}{Signal and channel}
			SISO, $L$-reflector IRS, $K$-user, $N$-subband
			\begin{itemize}
				\item transmit signal at the AP:
				\begin{equation}
					x(t)=\Re\left\{\sum_{n=1}^N\left({w_{I,n}\tilde{x}_{I,n}(t)}+w_{P,n}\right){e^{j2{\pi}{f_n}{t}}}\right\}
				\end{equation}
				\begin{itemize}
					\item $w_{I/P,n}$ collects magnitude and phase of the information and power signal
					\item $\tilde{x}_{I,n}\sim\mathcal{CN}(0,1)$ is the information symbol
				\end{itemize}
				\item composite channel for user $k$:
				\begin{itemize}
					\item direct: AP-user ($h_{D,k,n}$)
					\item extra: AP-IRS ($\boldsymbol{h}_{I,n}$), \alert{IRS reflection ($\boldsymbol{\Theta}$)}, IRS-user ($\boldsymbol{h}_{R,k,n}^H$)
					\begin{equation}
						h_{E,k,n} = \boldsymbol{h}_{R,k,n}^H \alert{\boldsymbol{\Theta}} \boldsymbol{h}_{I,n} = \boldsymbol{v}_{k,n}^H \alert{\boldsymbol{\phi}}
					\end{equation}
					\item sum up and stack over all subbands:
					\begin{equation}
						\boldsymbol{h}_k = \boldsymbol{h}_{D,k} + \boldsymbol{V}_k^H \alert{\boldsymbol{\phi}}
					\end{equation}
				\end{itemize}
				\item received signal by user $k$
				\begin{equation}
					y_k(t)=\Re\left\{\sum_{n=1}^N{h_{{k,n}}}\left({w_{I,n}\tilde{x}_{I,n}(t)}+w_{P,n}\right){e^{j2{\pi}{f_n}{t}}}\right\}
				\end{equation}
			\end{itemize}
		\end{frame}
	\end{subsection}

	\begin{subsection}{Information decoder}
		\begin{frame}{Information decoder}
			The power component $y_{P,k}(t)$ creates no interference to the information component $y_{I,k}(t)$. Hence, the achievable rate of user $k$ is
			\begin{equation}\label{eq:R_k}
				R_k(\boldsymbol{w}_I,\boldsymbol{\phi},\rho,\boldsymbol{\alpha}_k)=\sum_{n=1}^N\alpha_{k,n}{\log_2\left(1+\frac{(1-\rho)\lvert h_{k,n}w_{I,n} \rvert^2}{\sigma_n^2}\right)}
			\end{equation}
			\begin{itemize}
				\item $\rho$ is the splitting ratio for energy harvesting
				\item $\alpha_{k,n}$ is the allocation indicator:
				\begin{equation}
					\alpha_{k,n} =
					\begin{cases}
						1, & \text{if subband } n \text{ is given to user } k \\
						0, & \text{otherwise}
					\end{cases}
				\end{equation}
				\item $\sigma_n^2$ is the variance of the noise at RF band and during RF-to-BB conversion
			\end{itemize}
		\end{frame}
	\end{subsection}

	\begin{subsection}{Energy harvester}
		\begin{frame}{Energy harvester}
			A truncated Taylor expansion of small signal model highlights the dependency of harvester output DC current on the received waveform:
			\begin{equation}
				z_k(\boldsymbol{w}_I,\boldsymbol{w}_P,\boldsymbol{\phi},\rho)=\sum_{i\,\text{even},i\ge2}^{n_0}{k_i}{\rho^{i/2}}{R_{\text{ant}}^{i/2}}{\mathcal{E}\left\{\mathcal{A}\left\{y_k(t)^i\right\}\right\}}
			\end{equation}
			Pick $n_0=4$ and we have:
			\begin{figure}
				\centering
				\includegraphics[width=\linewidth]{assets/z.eps}
			\end{figure}
		\end{frame}
	\end{subsection}

	\begin{subsection}{Rate-energy region}
		\begin{frame}{Rate-energy region}
			Define the achievable weighted sum \textbf{R-E region} as
			\begin{equation}
				\begin{split}
					C_{R-I}(P)
					&\triangleq \biggl\{(R,Z):R\le\sum_{k=1}^K{u_{I,k}R_k},Z\le\sum_{k=1}^K u_{P,k}z_k,\\
					&\quad \frac{1}{2}({\boldsymbol{w}_I^H}{\boldsymbol{w}_I}+{\boldsymbol{w}_P^H}{\boldsymbol{w}_P}) \le P\biggr\}
				\end{split}
			\end{equation}
			\begin{itemize}
				\item $P$ is the transmit power budget
				\item $u_{I/P,k}$ are the information and power weight of user $k$
			\end{itemize}
		\end{frame}
	\end{subsection}
\end{section}

\begin{section}{Single-user optimization}
	\begin{subsection}{Problem formulation}
		\begin{frame}{Problem formulation}
			We characterize the rate-energy region through a current maximization problem subject to transmit power, rate, and IRS constraints
			\begin{maxi!}
				{\boldsymbol{w}_I,\boldsymbol{w}_P,\boldsymbol{\phi},\rho}{z(\boldsymbol{w}_I,\boldsymbol{w}_P,\boldsymbol{\phi},\rho)}{\label{op:su}}{\label{eq:su_target}}
				\addConstraint{\frac{1}{2}({\boldsymbol{w}_I^H}{\boldsymbol{w}_I}+{\boldsymbol{w}_P^H}{\boldsymbol{w}_P})\le{P}}
				\addConstraint{\sum_{n}{\log_2\left(1+\frac{(1-\rho)\lvert(h_{D,n}+\boldsymbol{v}_n^H\boldsymbol{\phi})w_{I,n}\rvert^2}{\sigma_n^2}\right)} \ge \bar{R}}\label{eq:su_rate_constraint}
				\addConstraint{\lvert{\phi_l}\rvert=1, \quad l=1,\dots,L}
				\addConstraint{0 \le \rho \le 1}
			\end{maxi!}
			\vspace{1em}
			$\boldsymbol{w}_{I/P},\boldsymbol{\phi},\rho$ are coupled in \ref{eq:su_target}, \ref{eq:su_rate_constraint}, consider \textbf{alternating optimization} (AO).
		\end{frame}
	\end{subsection}

	\begin{subsection}{IRS phase shift}
		\begin{frame}{Frequency-selective IRS}
			The ideal \textbf{Frequency-Selective} (FS) IRS provides:
			\begin{itemize}
				\item subband-dependent reflection coefficients ($\boldsymbol{\phi}_n$ replaces $\boldsymbol{\phi}$)
				\item total DoF $NL$
			\end{itemize}
			\vspace{1em}
			Note that $\lvert{(h_{D,n}+\boldsymbol{v}_n^H\boldsymbol{\phi}_n)w_{I,n}}\rvert \le \lvert{h_{D,n}w_{I,n}}\rvert+\lvert{\boldsymbol{v}_n^H\boldsymbol{\phi}_n w_{I,n}}\rvert$ and the equality holds when the direct and IRS-aided links are aligned. Therefore, we simply select the phase shift of element $l$ at subband $n$ as
			\begin{equation}
				\theta_{n,l}^\star = \angle{h}_{D,n} - \angle{h_{R,n,l}}-\angle{h_{I,n,l}}
			\end{equation}
			That is to say, the optimal phase shift for FS-IRS is obtained in closed form in the single-user scenario, which ensures the direct, extra and composite channels have the same phase.
		\end{frame}

		\begin{frame}{Frequency-flat IRS (1)}
			\textbf{Frequency-Flat} (FF) IRS reflects all subbands equally with a DoF of $L$.
			\begin{block}{Semi-Definite Relaxation (SDR): rate constraint}
				To simplify rate constraint \ref{eq:su_rate_constraint}, we observe that
				\begin{equation}
					\begin{split}
						\lvert{h_{D,n}+\boldsymbol{v}_n^H\boldsymbol{\phi}}\rvert^2
						&=\lvert{h_{D,n}}\rvert^2+h_{D,n}^*\boldsymbol{v}_n^H\boldsymbol{\phi}+\boldsymbol{\phi}^H\boldsymbol{v}_n{h_{D,n}}+\boldsymbol{\phi}^H\boldsymbol{v}\boldsymbol{v}^H\boldsymbol{\phi}\\
						&=\bar{\boldsymbol{\phi}}^H\boldsymbol{R}_n\bar{\boldsymbol{\phi}}=\mathrm{Tr}(\boldsymbol{R}_n\bar{\boldsymbol{\phi}}\bar{\boldsymbol{\phi}}^H)=\mathrm{Tr}(\boldsymbol{R}_n\boldsymbol{\Phi})
					\end{split}
				\end{equation}
				where $t$ is an auxiliary variable with unit modulus and
				\begin{equation}\label{eq:R_n,phi}
					\boldsymbol{R}_n=
					\begin{bmatrix}
						\boldsymbol{v}_n\boldsymbol{v}_n^H & \boldsymbol{v}_n{h_{D,n}} \\
						h_{D,n}^*{\boldsymbol{v}_n^H}      & h_{D,n}^*{h_{D,n}}
					\end{bmatrix},
					\quad \bar{\boldsymbol{\phi}}=\
					\begin{bmatrix}
						\boldsymbol{\phi} \\
						t
					\end{bmatrix},
					\quad \boldsymbol{\Phi}=\bar{\boldsymbol{\phi}}\bar{\boldsymbol{\phi}}^H
				\end{equation}
			\end{block}
		\end{frame}

		\begin{frame}{Frequency-flat IRS (2)}
			We optimize $\boldsymbol{\Phi}$ with given waveform $\boldsymbol{w}_{I/P}$ and splitting ratio $\rho$.
			\begin{block}{SDR: current expression}
				Define $\boldsymbol{M}=[\boldsymbol{V}^H,\boldsymbol{h}_D]^H$ such that $\boldsymbol{h}\boldsymbol{h}^H=\boldsymbol{M}^H\boldsymbol{\Phi}\boldsymbol{M}$. Introduce auxiliary variables
				\begin{equation}\label{eq:t}
					\begin{split}
						t_{I/P,n}
						&=\boldsymbol{h}^H\boldsymbol{W}_{I/P,n}^*\boldsymbol{h}=\mathrm{Tr}(\boldsymbol{h}\boldsymbol{h}^H\boldsymbol{W}_{I/P,n}^*)=\mathrm{Tr}(\boldsymbol{M}^H\boldsymbol{\Phi}\boldsymbol{M}\boldsymbol{W}_{I/P,n}^*)\\
						&=\mathrm{Tr}(\boldsymbol{M}\boldsymbol{W}_{I/P,n}^*\boldsymbol{M}^H\boldsymbol{\Phi})=\mathrm{Tr}(\boldsymbol{C}_{I/P,n}\boldsymbol{\Phi})
					\end{split}
				\end{equation}
				Therefore, the current expression rewrites as
				\begin{equation}\label{eq:z_irs}
					z(\boldsymbol{\Phi})=\frac{1}{2}{\beta_2}{\rho}(t_{I,0}+t_{P,0})+\frac{3}{8}{\beta_4}{\rho^2} \left(2t_{I,0}^2 + \sum_{n=-N+1}^{N-1}{t_{P,n}t_{P,n}^*}\right)+\frac{3}{2}{\beta_4}{\rho^2}t_{I,0}t_{P,0}
				\end{equation}
			\end{block}
		\end{frame}

		\begin{frame}{Frequency-flat IRS (3)}
			We use first-order Taylor expansion to approximate the second-order terms in \ref{eq:z_irs}. Based on the variables optimized at iteration $i - 1$, the local approximation at iteration $i$ gives
			\begin{align}
				(t_{I,0}^{(i)})^2
				& \ge 2 t_{I,0}^{(i)}t_{I,0}^{(i-1)} - (t_{I,0}^{(i-1)})^2\label{eq:t_{I,0}_square}\\
				t_{P,n}^{(i)} (t_{P,n}^{(i)})^*
				& \ge 2 \Re\left\{t_{P,n}^{(i)} (t_{P,n}^{(i-1)})^*\right\} - t_{P,n}^{(i-1)} (t_{P,n}^{(i-1)})^*\\
				t_{I,0}^{(i)} t_{P,0}^{(i)}
				& = \frac{1}{4}(t_{I,0}^{(i)} + t_{P,0}^{(i)})^2 - \frac{1}{4}(t_{I,0}^{(i)} - t_{P,0}^{(i)})^2\nonumber\\
				& \ge \frac{1}{2}(t_{I,0}^{(i)} + t_{P,0}^{(i)})(t_{I,0}^{(i-1)} + t_{P,0}^{(i-1)})\nonumber\\
				& \quad - \frac{1}{4}(t_{I,0}^{(i-1)} + t_{P,0}^{(i-1)})^2 - \frac{1}{4}(t_{I,0}^{(i)} - t_{P,0}^{(i)})^2\label{eq:t_{I/P,0}}
			\end{align}
		\end{frame}

		\begin{frame}{Frequency-flat IRS (4)}
			Hence, problem \ref{op:su} is transformed to
			\begin{maxi!}
				{\boldsymbol{\boldsymbol{\Phi}}}{\tilde{z}(\boldsymbol{\Phi})}{\label{op:su_irs}}{\label{eq:su_irs_target}}
				\addConstraint{\sum_{n}{\log_2\left(1+\frac{(1-\rho)\lvert{w}_{I,n}\rvert^2\mathrm{Tr}(\boldsymbol{R}_n\boldsymbol{\Phi})}{\sigma_n^2}\right)} \ge \bar{R}}
				\addConstraint{\boldsymbol{\Phi}_{l,l}=1, \quad l=1,\dots,L+1}
				\addConstraint{\boldsymbol{\Phi}\succeq{0}}
				\addConstraint{\mathrm{rank}(\boldsymbol{\Phi})=1\label{co:irs_rank}}
			\end{maxi!}
		\end{frame}
	\end{subsection}

	\begin{subsection}{Waveform}
		\begin{frame}{Waveform (1)}
			We optimize $\boldsymbol{w}_{I/P}$ with given IRS $\boldsymbol{\phi}$ and splitting ratio $\rho$.
			\begin{block}{SDR: current expression}
				Introduce auxiliary variables
				\begin{equation}\label{eq:t'}
					t_{I/P,n}' = \boldsymbol{w}_{I/P}^H \boldsymbol{H}_n^* \boldsymbol{w}_{I/P} = \mathrm{Tr}(\boldsymbol{H}_n^*\boldsymbol{W}_{I/P})
				\end{equation}
				Therefore, the current expression rewrites as
				\begin{equation}\label{eq:z_waveform}
					z(\boldsymbol{W}_{I/P})=\frac{1}{2} \beta_2 \rho (t'_{I,0}+t'_{P,0})+\frac{3}{8} \beta_4 \rho^2 \left(2(t'_{I,0})^2 + \sum_{n=-N+1}^{N-1}{t'_{P,n}(t'_{P,n})^*}\right)+\frac{3}{2} \beta_4 \rho^2 t'_{I,0}t'_{P,0}
				\end{equation}
				Since \ref{eq:z_waveform} and \ref{eq:z_irs} are in the same form, we use same Taylor approximation.
			\end{block}
		\end{frame}

		\begin{frame}{Waveform (2)}
			Hence, problem \ref{op:su} is transformed to
			\begin{maxi!}
				{\boldsymbol{W}_{I/P}}{\tilde{z}(\boldsymbol{W}_I,\boldsymbol{W}_P)}{\label{op:su_waveform}}{\label{eq:su_waveform_target}}
				\addConstraint{\sum_{n}{\log_2\left(1+\frac{(1-\rho)W_{I,n,n}\lvert h_n \rvert^2}{\sigma_n^2}\right)} \ge \bar{R}\label{eq:su_waveform_constraint}}
				\addConstraint{\frac{1}{2}\left(\mathrm{Tr}(\boldsymbol{W}_I)+\mathrm{Tr}(\boldsymbol{W}_P)\right) \le P}
				\addConstraint{\boldsymbol{W}_{I/P} \succeq 0}
				\addConstraint{\mathrm{rank}(\boldsymbol{W}_{I/P})=1\label{co:waveform_rank}}
			\end{maxi!}
		\end{frame}
	\end{subsection}

	\begin{subsection}{Splitting ratio}
		\begin{frame}{Splitting ratio}
			We optimize $\rho$ with given IRS $\boldsymbol{\phi}$ and waveform $\boldsymbol{w}_{I/P}$.
			\begin{maxi!}
				{\rho}{z(\rho)}{\label{op:su_ratio}}{\label{eq:su_ratio_target}}
				\addConstraint{\sum_{n}{\log_2\left(1+\frac{(1-\rho)\lvert h_n w_n \rvert^2}{\sigma_n^2}\right)} \ge \bar{R}\label{eq:su_ratio_constraint}}
				\addConstraint{0 \le \rho \le 1}
			\end{maxi!}
			Since $z(\rho)$ is a quadratic function that monotonically increases over $\rho \in [0, 1]$, we replace $z(\rho)$ with affine $\rho$ and solve problem \ref{op:su_ratio}.
		\end{frame}
	\end{subsection}
\end{section}
